A Review of Neutron Imaging

### 2020-10-15 neutron-imaging/neutron.tex
Final version.

Its mean square radius is $0.8 \times 10^{-15}$ m
\cite{bogdan2015particles}. It has no electric charge. However, it has
an intrinsic spin of $1/2$ (in units of $\hslash$)
\cite{basdevant2005fundamentals}. As a result, although it is unaffected by 
an electric field, it does respond to a magnetic field. In fact, a neutron 
responds to all four fundamental forces in Nature. Outside of
the nucleus, the neutron is unstable. It decays into a proton, an electron and
an electron neutrino with a half life of about $10$ minutes and $10$ seconds
\cite{Nakamura_2010}. The energy of decay is $0.782$ MeV \cite{heyde2004basic}.
A small fraction, about $0.001$, of free neutrons also produce gamma radiation 
while decaying. An even smaller fraction, about $4$ per million, of neutrons
produce electrons with an energy lesser than $13.6$ eV. As a result, they
stay stay bound to the proton forming a hydrogen atom. The fact that neutrons
decay cannot be ignored while using them for imaging. This feature of the
neutrons distingushes them from the other probes used for imaging like optical
photons, X-rays, gamma rays, infra-red radiation, sound waves and electrons. 
The latter probes are not known to decay.

@@

One can go back to
Chadwick's experiment \cite{chadwick1932existence} to devise yet another source
of neutrons. Chadwick bombarded a slab of Beryllium with alpha radiation 
emitted from decaying Polonium nuclei. The nuclear reaction
\begin{equation}\label{s3e1}
{}_2^4\text{He} + {}_4^9\text{Be} \rightarrow {}_6^{12}\text{C}+{}_0^1\text{n}
\end{equation}
produced the first neutron detected in a laboratory. Some other reactions that 
also yeild neutrons are\cite{turner2008atoms}
\begin{eqnarray}
{}^2_1\text{H} + {}^3_1\text{H}&\rightarrow&{}_2^{4}\text{He}+{}_0^1\text{n}
\label{s3e2} \\
{}^2_1\text{H} + {}^2_1\text{H}&\rightarrow&{}_2^{3}\text{He}+{}_0^1\text{n}
\label{s3e3} \\
{}^2_1\text{H} + {}_6^{12}\text{C}&\rightarrow&{}_7^{13}\text{N}+{}_0^1\text{n}
\label{s3e4} \\
{}^1_1\text{p} + {}_3^1\text{H}&\rightarrow&{}_2^{4}\text{He}+{}_0^1\text{n}
\label{s3e5} \\
{}_1^1\text{p} + {}_3^7\text{Li}&\rightarrow&{}_4^{7}\text{Be}+{}_0^1\text{n}
 \label{s3e6} 
\end{eqnarray}
Each of these processes produce neutrons with a continuous energy spectrum. If
we need neutrons with narrow energy spread then we have to consider nuclear
reactions triggered by photons. Chadwick and Goldhaber 
\cite{chadwick1934nuclear} bombarded Deuterium nuclei with gamma rays of energy
exceeding $2.22$ MeV to trigger the reaction
\begin{equation}\label{s3e7}
{}^2_1\text{H} + \gamma \rightarrow {}^1_1 \text{H} + {}^1_0 n.
\end{equation}
Similarly, a photon of energy more than $1.67$ MeV can disintegrate Beryllium
into two Helium nuclei and a neutron.
\begin{equation}\label{s3e8}
{}^9_4\text{Be} + \gamma \rightarrow 2 {}^4_2\text{He} + {}^1_0 n.
\end{equation}
The $\gamma$-rays themselves come from radioactive ${}^{124}_{51}$Sb which 
spontaneously disintegrates as
\begin{equation}\label{s3e9}
{}^{124}_{51}\text{Sb} \rightarrow {}^{124}_{52}\text{Te} + e^{-1} + \gamma.
\end{equation}
Other commonly used isotopes to generate $\gamma$ radiation are ${}^{24}_{11}$
Na, ${}^{116}_{49}$In, ${}^{140}_{57}$La and ${}^{226}_{88}$Ra. A recent method
of generating neutrons involes spallation of heavy nuclei due to impact of 
very high energy protons. The term spallation means ejection of fragments from
a body due to extraordinarily large stress. In the case of nuclear spallation,
$\text{H}^{-}$ ions are accelerated to energy up to $1$ GeV before passing
them through a foil. The latter process strips the $\text{H}^{-}$ ion of its
two electrons and leaves behind a bare proton. The protons so generated are
made to pass at relativistic speeds into an accumulator ring. They are 
ejected out of the ring at a $60$ Hz into a tank of liquid Hg where they
trigger spallation of Hg nuclei. The greatest benefit of spallation is that
the resulting neutrons do not trigger a chain reaction as they do in the case
of nuclear fission reactors.

@@

\section{Detection of neutrons}\label{s5}
Neutrons are chargeless and their magnetic moment is too low to be used for
reliably detecting their presence. They also have a small electric dipole 
moment but it has not been measured in a laboratory. Free neutrons decay to
produce a proton and an electron but its half life of around $10$ minutes is
too long to permit the detection of neutrons through its decay products. That
leaves us with just one option, consider nuclear reactions triggered by the
neutrons and measure their results to estimate the number and energy of the
neutrons. Since the imaging applications involve mostly slow neutrons, we will
focus our attention on their detection.

@@

Bubble detectors are inexpensive and portable neutron detecting devices. They
consist of a transparent polymer with a large number of tiny droplets of 
a superheated liquid. The passage of a charged particle through such droplets
results in a rapid phase change from liquid to vapor. The transparent droplets
become visible bubbles which are counted either manually or electronically.
\section{Neutron dosimetry and shielding}\label{s6}
Radiation Dosimetry is the measurement of the `dose' of radiation absorbed by
a biological sample. Dosimetry is important for the safety of people working
with radiation. Biological materials are rich in protons. When they are
exposed to neutrons, the elastic collision of the neutrons and the protons
leads to a transfer of a substantial amount of energy to the protons resulting
in breaking of chemical bonds. Neutrons damage biological samples more than
a comparable dose of electrons or $\gamma$ rays. The amount of energy absorbed
by $1$ kg of a material is called the `absorbed dose'. It is measured in grays,
where $1$ gray is equal to $1$ Jkg${}^{-1}$. It is measured and specified at 
every point of the exposed sample. Before getting into the details of neutron
dosimetry, we introduce a few physical concepts.

@@

A neutron source produces neutrons with a wide range of wavelengths. A device
that allows us to select one among them is called a monochromator. Crystal
monochromators are based on Bragg's law
\begin{equation}\label{s7e2}
n\lambda = 2d\sin\theta,
\end{equation}
where $n$ is the `order' of diffraction, $\lambda$ is the wavelength of 
neutrons, $d$ is the distance between the lattice planes and $\theta$ is the
angle between the crystal and the incident beam. (The angle of incidence is
$\pi/2 - \theta$.) Once a crystal is chosen, $d$ is fixed. One can select
the wavelength of choice by changing $\theta$. The intensity of the diffracted
beam reduces for higher order of diffractions. Therefore, monochromators us
the first order beam.

@@

The existence of spin states of neutrons allows us to describe beams of 
neutrons by their polarization states. Being a spin-half particle, a neutron
can exist in one of $|+\rangle$ of $|-\rangle$ states. If a beam has $N_+$
neutrons in $|+\rangle$ and $N_-$ neutrons in $|-\rangle$ states then its
polarization is defined to be
\begin{equation}\label{s7e5}
P = \frac{N_+ - N-}{N_+ + N_-}.
\end{equation}
It is equally common to express beam polarization in terms of flipping ratio
as
\begin{equation}\label{s7e6}
F = \frac{N_+}{N_-} = \frac{1 - P}{1 + P}.
\end{equation}
When the nuclear and the magnetic structure factors of a crystal are equal in
magnitude, the total structure factor vanishes for one of the two spin states.
As a result, the Bragg reflection from the crystal plance produces a beam with
almost all neutrons in the remaining polarization states. The most commonly
used crystal to achieve neutron polarization is the [111] reflection of 
Heusler alloy \cite{anderson2008neutron}. This technique is most suitable for
steady state source of neutrons \cite{williams1988polarized}. In the case of
a pulsed beam of neutrons it is common to use resonance absorption filters.
When an unpolarized neutron beam whose energy corresponds to a nuclear
resonance is shone on a target of polarized nuclear spins, one of the 
neutron spins is absorbed while the other one is transmitted 
\cite{williams1988polarized}.

@@

The intensity of the incident and the transmitted beams is given by equation
\eqref{s4e2}. If $I_0$ and $I$ are the intensities of the incident and the
transmitted beams then the degree of darkness of the film used to visualize
the image is
\begin{equation}\label{s8e1}
D_e = \ln\frac{I_0}{I}.
\end{equation}
The film brightens most where the transmitted intensity is the highest. In the
case of a digital receptor, the grey level value $G$ is
\begin{equation}\label{s8e2}
G = Cn + G_0,
\end{equation}
where $C$ is a constant, $n$ is the neutron density and $G_0$ is another 
constant that indicates the gray level due to noise in the circuits. The 
presence of $G_0$ suggests that $G \ne 0$ even in the absence of neutron flux.
This method suffers from the disadvantage of catching signals not only from
the transmitted neutrons but from other ambient radiation as well. As a result,
the images often are blurred. Nevertheless, it is used in situation where the
sharpness of the image is not critical. For example, in the non-destructive
testing in materials, it sufficient to know if the material has contaminants,
defects, lubrication etc. It is also used in the investigation of two-phase
flows, flows through porous media \cite{anderson2008neutron}, detection of
hydrgenous materials and moisture \cite{richards2004neutron}.

\section{Neutron tomography}\label{s9}
Tomography is an imaging technique in which the image of a target is constructed
by combining the images of its slices. The most usage is in medical imaging
in the form of CT (Computerized Tomography) scans. X-rays are used over a 
large number of slices of the human body. The `image' produced by each slice
is combined with others to get complete information of the portion of interest.
The way 2-d neutron imaging is an adaptation of 2-d X-ray imaging, neutron
tomography is applying the principles of X-ray tomography to neutron imaging. 
Consider a target $T$ whose image we are are interested in. We place the origin
of the cartesian coordinate system at a suitable point in $T$. We consider $T$
to be composed of a large number of slices whose normal is along the $z$-axis.
We know from equation \eqref{s4e2} that the attenuation of intensity of an
incident neutron beams by a slice $S$ is described by the equation
\begin{equation}\label{s9e1}
I = I_0\exp\left(-\int_S\mu(x, y)ds\right).
\end{equation}
Since the target is not a homogenous object, we cannot consider it to have a
uniform attenuation $\mu$ as we did in equation \eqref{s4e2}. The parameter
$s$ is the coordinate along the direction of propagation of neutrons in the
target. We can rearrange equation \eqref{s9e1} and define a function
\begin{equation}\label{s9e2}
p(t, \theta) = \ln\left(\frac{I}{I_0}\right) = -\int_S\mu(x, y)ds.
\end{equation}
The coordinate $t$ runs perpendicular to the coordinate $s$. One can write 
this in terms of an integral that extends over all $s$ as
\begin{equation}\label{s9e3}
p(t, \theta) = 
-\int_{-\infty}^\infty \mu(x, y)\delta(t - x\cos\theta - y\sin\theta)ds.
\end{equation}
The function $p(t, \theta)$ is called the Radon transform of the slice. It is
taken along the line $t = x\cos\theta + y\sin\theta$, expressed in the 
normal form \cite{gonzales1992digital}. It is the projection of the function
$f$ on the line $t = x\cos\theta + y\sin\theta$. It can be shown using the
Fourier Slice Theorem \cite{bracewell1956strip} that a Fourier transform of
$p(t, \theta)$ is the same as the slice of the 2-d Fourier transform of the
target. Thus, taking the 2-d inverse Fourier transform of the result of the
first operation will give us the original function. The number of such
projections is determined by Shannon's sampling theorem.

@@

Phase differences can either be recorded as a 2-d image or a tomogram. In the
latter case, the 3-d distribution of the refractive index of the target is 
reconstructed. The phase differences can either be represented as the 
differences in their values or the gradient of phase shift or the laplacian of
the phase shift \cite{anderson2008neutron}.

\section{Larmor precession based imaging}\label{s11}
A magnetic dipole of moment $\vec{\mu}$ experiences a torque $\vec{\tau} = 
\vec{\mu} \times \vec{B}$ in a magnetic field $\vec{B}$. If $\gamma$ is the
gyromagnetic ratio of the dipole then $\vec{\mu} = \gamma\vec{L}$, $\vec{L}$
being its angular momentum. Since
\begin{equation}\label{s11e1}
\frac{d\vec{L}}{dt} = \vec{\tau} = \gamma\vec{L} \times \vec{B},
\end{equation}
when $\vec{B}$ is not parallel (or anti-parallel) to $\vec{L}$, the vector 
$\vec{L}$ precesses about $\vec{B}$ with a frequency $\omega = \gamma B$, 
called the Larmor frequency. When a polarized beam of neutrons is made to
pass through a magnetic material, the spin vector of the neutrons undergoes
Larmor precession in the internal magnetic field of material. The angle of 
spin changes by an amount that depends on the magnitude of the magnetic field
and the time the neutron was present in it. The change in the spin orientation
can be measured by a spin analyzer. Different regions of the material show
different changes in spin orientation. By mapping the change in spin to grey
scale one can create a contrast between these regions 
\cite{kardjilov2011neutron}. This is the principle behind Larmor precession
based neutron imaging. If the magnetic fields are sufficiently low so that
the Larmor tilt is never more than $\pm\pi$, quantitative 2-d and 3-d imaging
is possible. It was used to study the trapping of magnetic flux in type-I 
superconductors \cite{kardjilov2008three}. This idea can be extended to 
examine the how a ferromagnet becomes a paramagnet past the Curie temperature.
When a beam of polarized neutrons falls on a ferromagnet, the randomly oriented
magnetic domains of the ferromagnet depolarize the beam. When the same material
is heated, the domains begin to weaken and beyond the Curie temperature they
disappear altogether. The polarized beam is not transmitted undisturbed. This
technique can be used to study the change in magnetic properties of materials
in response to changing stress, pressure or applied electromagnetic fields
\cite{kardjilov2011neutron}.

\section{Spin-echo based imaging}\label{s12}
Consider a beam of polarized nucleons (neutrons, protons, or nuclei with an odd
number of nucleons) passes through a region of homogenous magnetic field its
spin vector tilts with respect to the original direction because of Larmor
precession. When the nucleons emerge from the region, they all emerge with the
same tilt. However, as time goes by, different nucleons have a different angle
of tilt. This happens because of natural spin relaxation and due to 
inhomogeneities in the surroundings of the nucleons. If the nucleons are 
subjected to a field that flips the spin by $\pi$ then after a while all spins
that had dispersed due to inhomogeneities once again orient themselves together
creating an `echo'. The magnitude of the echo is not the same as that of the
original beam because of the natural spin relaxation. This is the principle
behind the 'spin-echo' phenomenon used in nuclear magnetic resonance 
spectroscopy and magnetic resonance imaging.

@@

\section{Imaging of biologically important structures}\label{s13}
Biological targets have an extremely rich mixture of molecules. A knowledge of
the presence of certain trace elements in the targets reveals important 
information about its functioning, health and pathology. Few imaging techniques
can accurately provide information about the elements. The chemical shift 
imaging in magnetic resonance imaging provides information at the molecular
level. It provides little information about the elements present. This is where
neutron imaging comes into the picture. When neutrons undergo an inelastic
scattering with the nuclei of the target, they nuclei get into an excited
state. They get back to their ground state after emitting a photon. We now 
have a detailed knowledge of the nuclear energy levels. By measuring the energy
of the emitted photon we can guess the nucleus, and thus the element, which
might have produced it. This is the principle behing the neutron stimulated
emission computed tomography (NSECT) imaging.
\section{Conclusion}\label{s14}
\bibliography{neutron}
\bibliographystyle{plain}
\end{document}

>>>>

Its mean square radius is $0.8 \times 10^{-15}$ m
\cite{bogdan2015particles}. It has no electric charge. However, it has
an intrinsic spin of $1/2$ (in units of $\hslash$)
\cite{basdevant2005fundamentals}. As a result, although it is unaffected by 
an electric field, it does respond to a magnetic field. In fact, a neutron 
responds to all four fundamental forces in Nature. Outside of
the nucleus, the neutron is unstable. It decays into a proton, an electron and
an electron neutrino with a half life of about $10$ minutes and $10$ seconds
\cite{Nakamura_2010}. The energy of decay is $0.782$ MeV \cite{heyde2004basic}.
A small fraction, about $0.001$, of free neutrons also produce gamma radiation 
while decaying. An even smaller fraction, about $4$ per million, of neutrons
produce electrons with an energy lesser than $13.6$ eV. As a result, they
stay stay bound to the proton forming a hydrogen atom. The fact that neutrons
decay cannot be ignored while using them for imaging. This feature of the
neutrons distinguishes them from the other probes used for imaging like optical
photons, X-rays, gamma rays, infra-red radiation, sound waves and electrons. 
The latter probes are not known to decay.

@@

One can go back to
Chadwick's experiment \cite{chadwick1932existence} to devise yet another source
of neutrons. Chadwick bombarded a slab of Beryllium with alpha radiation 
emitted from decaying Polonium nuclei. The nuclear reaction
\begin{equation}\label{s3e1}
{}_2^4\text{He} + {}_4^9\text{Be} \rightarrow {}_6^{12}\text{C}+{}_0^1\text{n}
\end{equation}
produced the first neutron detected in a laboratory. Some other reactions that 
also yield neutrons are\cite{turner2008atoms}
\begin{eqnarray}
{}^2_1\text{H} + {}^3_1\text{H}&\rightarrow&{}_2^{4}\text{He}+{}_0^1\text{n}
\label{s3e2} \\
{}^2_1\text{H} + {}^2_1\text{H}&\rightarrow&{}_2^{3}\text{He}+{}_0^1\text{n}
\label{s3e3} \\
{}^2_1\text{H} + {}_6^{12}\text{C}&\rightarrow&{}_7^{13}\text{N}+{}_0^1\text{n}
\label{s3e4} \\
{}^1_1\text{p} + {}_3^1\text{H}&\rightarrow&{}_2^{4}\text{He}+{}_0^1\text{n}
\label{s3e5} \\
{}_1^1\text{p} + {}_3^7\text{Li}&\rightarrow&{}_4^{7}\text{Be}+{}_0^1\text{n}
 \label{s3e6} 
\end{eqnarray}
Each of these processes produce neutrons with a continuous energy spectrum. If
we need neutrons with narrow energy spread then we have to consider nuclear
reactions triggered by photons. Chadwick and Goldhaber 
\cite{chadwick1934nuclear} bombarded Deuterium nuclei with gamma rays of energy
exceeding $2.22$ MeV to trigger the reaction
\begin{equation}\label{s3e7}
{}^2_1\text{H} + \gamma \rightarrow {}^1_1 \text{H} + {}^1_0 n.
\end{equation}
Similarly, a photon of energy more than $1.67$ MeV can disintegrate Beryllium
into two Helium nuclei and a neutron.
\begin{equation}\label{s3e8}
{}^9_4\text{Be} + \gamma \rightarrow 2 {}^4_2\text{He} + {}^1_0 n.
\end{equation}
The $\gamma$-rays themselves come from radioactive ${}^{124}_{51}$Sb which 
spontaneously disintegrates as
\begin{equation}\label{s3e9}
{}^{124}_{51}\text{Sb} \rightarrow {}^{124}_{52}\text{Te} + e^{-1} + \gamma.
\end{equation}
Other commonly used isotopes to generate $\gamma$ radiation are ${}^{24}_{11}$
Na, ${}^{116}_{49}$In, ${}^{140}_{57}$La and ${}^{226}_{88}$Ra. A recent method
of generating neutrons involves spallation of heavy nuclei due to impact of 
very high energy protons. The term spallation means ejection of fragments from
a body due to extraordinarily large stress. In the case of nuclear spallation,
$\text{H}^{-}$ ions are accelerated to energy up to $1$ GeV before passing
them through a foil. The latter process strips the $\text{H}^{-}$ ion of its
two electrons and leaves behind a bare proton. The protons so generated are
made to pass at relativistic speeds into an accumulator ring. They are 
ejected out of the ring at a $60$ Hz into a tank of liquid Hg where they
trigger spallation of Hg nuclei. The greatest benefit of spallation is that
the resulting neutrons do not trigger a chain reaction as they do in the case
of nuclear fission reactors.

@@

\section{Detection of neutrons}\label{s5}
Neutrons are charge less and their magnetic moment is too low to be used for
reliably detecting their presence. They also have a small electric dipole 
moment but it has not been measured in a laboratory. Free neutrons decay to
produce a proton and an electron but its half life of around $10$ minutes is
too long to permit the detection of neutrons through its decay products. That
leaves us with just one option, consider nuclear reactions triggered by the
neutrons and measure their results to estimate the number and energy of the
neutrons. Since the imaging applications involve mostly slow neutrons, we will
focus our attention on their detection.

@@

Bubble detectors are inexpensive and portable neutron detecting devices. They
consist of a transparent polymer with a large number of tiny droplets of 
a super-heated liquid. The passage of a charged particle through such droplets
results in a rapid phase change from liquid to vapor. The transparent droplets
become visible bubbles which are counted either manually or electronically.
\section{Neutron dosimetry and shielding}\label{s6}
Radiation Dosimetry is the measurement of the `dose' of radiation absorbed by
a biological sample. Dosimetry is important for the safety of people working
with radiation. Biological materials are rich in protons. When they are
exposed to neutrons, the elastic collision of the neutrons and the protons
leads to a transfer of a substantial amount of energy to the protons resulting
in breaking of chemical bonds. Neutrons damage biological samples more than
a comparable dose of electrons or $\gamma$ rays. The amount of energy absorbed
by $1$ kg of a material is called the `absorbed dose'. It is measured in grays,
where $1$ gray is equal to $1$ Jkg${}^{-1}$. It is measured and specified at 
every point of the exposed sample. Before getting into the details of neutron
dosimetry, we introduce a few physical concepts.

@@

A neutron source produces neutrons with a wide range of wavelengths. A device
that allows us to select one among them is called a monochromator. Crystal
monochromators are based on Bragg's law
\begin{equation}\label{s7e2}
n\lambda = 2d\sin\theta,
\end{equation}
where $n$ is the `order' of diffraction, $\lambda$ is the wavelength of 
neutrons, $d$ is the distance between the lattice planes and $\theta$ is the
angle between the crystal and the incident beam. (The angle of incidence is
$\pi/2 - \theta$.) Once a crystal is chosen, $d$ is fixed. One can select
the wavelength of choice by changing $\theta$. The intensity of the diffracted
beam reduces for higher order of diffraction. Therefore, monochromators us
the first order beam.

@@

The existence of spin states of neutrons allows us to describe beams of 
neutrons by their polarization states. Being a spin-half particle, a neutron
can exist in one of $|+\rangle$ of $|-\rangle$ states. If a beam has $N_+$
neutrons in $|+\rangle$ and $N_-$ neutrons in $|-\rangle$ states then its
polarization is defined to be
\begin{equation}\label{s7e5}
P = \frac{N_+ - N-}{N_+ + N_-}.
\end{equation}
It is equally common to express beam polarization in terms of flipping ratio
as
\begin{equation}\label{s7e6}
F = \frac{N_+}{N_-} = \frac{1 - P}{1 + P}.
\end{equation}
When the nuclear and the magnetic structure factors of a crystal are equal in
magnitude, the total structure factor vanishes for one of the two spin states.
As a result, the Bragg reflection from the crystal plane produces a beam with
almost all neutrons in the remaining polarization states. The most commonly
used crystal to achieve neutron polarization is the [111] reflection of 
Heusler alloy \cite{anderson2008neutron}. This technique is most suitable for
steady state source of neutrons \cite{williams1988polarized}. In the case of
a pulsed beam of neutrons it is common to use resonance absorption filters.
When an unpolarized neutron beam whose energy corresponds to a nuclear
resonance is shone on a target of polarized nuclear spins, one of the 
neutron spins is absorbed while the other one is transmitted 
\cite{williams1988polarized}.

@@

The intensity of the incident and the transmitted beams is given by equation
\eqref{s4e2}. If $I_0$ and $I$ are the intensities of the incident and the
transmitted beams then the degree of darkness of the film used to visualize
the image is
\begin{equation}\label{s8e1}
D_e = \ln\frac{I_0}{I}.
\end{equation}
The film brightens most where the transmitted intensity is the highest. In the
case of a digital receptor, the Grey level value $G$ is
\begin{equation}\label{s8e2}
G = Cn + G_0,
\end{equation}
where $C$ is a constant, $n$ is the neutron density and $G_0$ is another 
constant that indicates the gray level due to noise in the circuits. The 
presence of $G_0$ suggests that $G \ne 0$ even in the absence of neutron flux.
This method suffers from the disadvantage of catching signals not only from
the transmitted neutrons but from other ambient radiation as well. As a result,
the images often are blurred. Nevertheless, it is used in situation where the
sharpness of the image is not critical. For example, in the non-destructive
testing in materials, it sufficient to know if the material has contaminants,
defects, lubrication etc. It is also used in the investigation of two-phase
flows, flows through porous media \cite{anderson2008neutron}, detection of
hydrogenous materials and moisture \cite{richards2004neutron}.

\section{Neutron tomography}\label{s9}
Tomography is an imaging technique in which the image of a target is constructed
by combining the images of its slices. The most usage is in medical imaging
in the form of CT (Computerized Tomography) scans. X-rays are used over a 
large number of slices of the human body. The `image' produced by each slice
is combined with others to get complete information of the portion of interest.
The way 2-d neutron imaging is an adaptation of 2-d X-ray imaging, neutron
tomography is applying the principles of X-ray tomography to neutron imaging. 
Consider a target $T$ whose image we are are interested in. We place the origin
of the Cartesian coordinate system at a suitable point in $T$. We consider $T$
to be composed of a large number of slices whose normal is along the $z$-axis.
We know from equation \eqref{s4e2} that the attenuation of intensity of an
incident neutron beams by a slice $S$ is described by the equation
\begin{equation}\label{s9e1}
I = I_0\exp\left(-\int_S\mu(x, y)ds\right).
\end{equation}
Since the target is not a homogeneous object, we cannot consider it to have a
uniform attenuation $\mu$ as we did in equation \eqref{s4e2}. The parameter
$s$ is the coordinate along the direction of propagation of neutrons in the
target. We can rearrange equation \eqref{s9e1} and define a function
\begin{equation}\label{s9e2}
p(t, \theta) = \ln\left(\frac{I}{I_0}\right) = -\int_S\mu(x, y)ds.
\end{equation}
The coordinate $t$ runs perpendicular to the coordinate $s$. One can write 
this in terms of an integral that extends over all $s$ as
\begin{equation}\label{s9e3}
p(t, \theta) = 
-\int_{-\infty}^\infty \mu(x, y)\delta(t - x\cos\theta - y\sin\theta)ds.
\end{equation}
The function $p(t, \theta)$ is called the Radon transform of the slice. It is
taken along the line $t = x\cos\theta + y\sin\theta$, expressed in the 
normal form \cite{gonzales1992digital}. It is the projection of the function
$f$ on the line $t = x\cos\theta + y\sin\theta$. It can be shown using the
Fourier Slice Theorem \cite{bracewell1956strip} that a Fourier transform of
$p(t, \theta)$ is the same as the slice of the 2-d Fourier transform of the
target. Thus, taking the 2-d inverse Fourier transform of the result of the
first operation will give us the original function. The number of such
projections is determined by Shannon's sampling theorem.

@@

Phase differences can either be recorded as a 2-d image or a tomogram. In the
latter case, the 3-d distribution of the refractive index of the target is 
reconstructed. The phase differences can either be represented as the 
differences in their values or the gradient of phase shift or the Laplacian of
the phase shift \cite{anderson2008neutron}.

\section{Larmor precession based imaging}\label{s11}
A magnetic dipole of moment $\vec{\mu}$ experiences a torque $\vec{\tau} = 
\vec{\mu} \times \vec{B}$ in a magnetic field $\vec{B}$. If $\gamma$ is the
gyro-magnetic ratio of the dipole then $\vec{\mu} = \gamma\vec{L}$, $\vec{L}$
being its angular momentum. Since
\begin{equation}\label{s11e1}
\frac{d\vec{L}}{dt} = \vec{\tau} = \gamma\vec{L} \times \vec{B},
\end{equation}
when $\vec{B}$ is not parallel (or anti-parallel) to $\vec{L}$, the vector 
$\vec{L}$ precesses about $\vec{B}$ with a frequency $\omega = \gamma B$, 
called the Larmor frequency. When a polarized beam of neutrons is made to
pass through a magnetic material, the spin vector of the neutrons undergoes
Larmor precession in the internal magnetic field of material. The angle of 
spin changes by an amount that depends on the magnitude of the magnetic field
and the time the neutron was present in it. The change in the spin orientation
can be measured by a spin analyzer. Different regions of the material show
different changes in spin orientation. By mapping the change in spin to Grey
scale one can create a contrast between these regions 
\cite{kardjilov2011neutron}. This is the principle behind Larmor precession
based neutron imaging. If the magnetic fields are sufficiently low so that
the Larmor tilt is never more than $\pm\pi$, quantitative 2-d and 3-d imaging
is possible. It was used to study the trapping of magnetic flux in type-I 
superconductors \cite{kardjilov2008three}. This idea can be extended to 
examine the how a ferromagnet becomes a paramagnet past the Curie temperature.
When a beam of polarized neutrons falls on a ferromagnet, the randomly oriented
magnetic domains of the ferromagnet depolarize the beam. When the same material
is heated, the domains begin to weaken and beyond the Curie temperature they
disappear altogether. The polarized beam is not transmitted undisturbed. This
technique can be used to study the change in magnetic properties of materials
in response to changing stress, pressure or applied electromagnetic fields
\cite{kardjilov2011neutron}.

\section{Spin-echo based imaging}\label{s12}
Consider a beam of polarized nucleons (neutrons, protons, or nuclei with an odd
number of nucleons) passes through a region of homogeneous magnetic field its
spin vector tilts with respect to the original direction because of Larmor
precession. When the nucleons emerge from the region, they all emerge with the
same tilt. However, as time goes by, different nucleons have a different angle
of tilt. This happens because of natural spin relaxation and due to 
inhomogeneities in the surroundings of the nucleons. If the nucleons are 
subjected to a field that flips the spin by $\pi$ then after a while all spins
that had dispersed due to inhomogeneities once again orient themselves together
creating an `echo'. The magnitude of the echo is not the same as that of the
original beam because of the natural spin relaxation. This is the principle
behind the 'spin-echo' phenomenon used in nuclear magnetic resonance 
spectroscopy and magnetic resonance imaging.

@@

\section{Imaging of biologically important structures}\label{s13}
Biological targets have an extremely rich mixture of molecules. A knowledge of
the presence of certain trace elements in the targets reveals important 
information about its functioning, health and pathology. Few imaging techniques
can accurately provide information about the elements. The chemical shift 
imaging in magnetic resonance imaging provides information at the molecular
level. It provides little information about the elements present. This is where
neutron imaging comes into the picture. When neutrons undergo an inelastic
scattering with the nuclei of the target, they nuclei get into an excited
state. They get back to their ground state after emitting a photon. We now 
have a detailed knowledge of the nuclear energy levels. By measuring the energy
of the emitted photon we can guess the nucleus, and thus the element, which
might have produced it. This is the principle behind the neutron stimulated
emission computed tomography (NSECT) imaging.

In the previous section we made a passing reference to small angle neutron
scattering as a technique complementary to neutron spin-echo imaging. We will
now briefly outline the small angle scattering technique. It is used to gather
structural and functional information about bio-molecules with a resolution
in the range $1$ to $100$ nm. Biological macro-molecules do not have the
structural order of a crystal. The scattering of neutrons by the nuclei of 
such molecules does not produce a sharp diffraction pattern seen with crystals.
Rather, they just produce an interference pattern. The intensity of the 
interference pattern depends on the scattering length density. Neutrons have
a different scattering length for Hydrogen and Deuterium nuclei and this fact
is used to generate contrast in the images. Either the entire molecule is 
imaged in a solution of D${}_2$O or a sub-unit of the molecule is marked with
Deuterium to get the contrast \cite{anderson2008neutron}.

\section{Conclusion}\label{s14}
In this article we reviewed the physics of neutrons from the perspective of its
role in probing materials at various scales of resolution. We started by 
mentioning the physical properties of neutrons, the way they can be produced,
how they can be detected and how a beam of neutrons can be manipulated. The
electrical neutrality of neutrons allows them to penetrate deeply in materials.
They also have a large collision cross-section with protons, making them
potentially impactful for biological targets. As a result, one must be careful
in exposing living matter to a beam of neutrons. We covered methods to measure
dose of neutron radiation and ways to shield neutrons. Optical and electron
microscopes rely heavily on our ability to manipulate light rays and beams of
electrons. The electrical neutrality of neutrons and their weak intrinsic
magnetic moment make it hard to manipulate their beams. Nevertheless, a few
techniques are available and we covered them in the section on optics.

In the second half of the article we reviewed imaging techniques starting with
the simplest 2-d imaging in which we just record the intensity of the 
transmitted beam and interpret it like an X-ray image. We then covered the
tomographic technique borrowed from X-rays and described its superiority over
2-d imaging in getting the distribution of nuclei in targets. Neutrons also 
have a complex refractive index. The imaginary part of the refractive index is
used to differentiate between nuclei that have similar attenuation properties.
The magnetic moment of neutrons permit them to be treated like protons in the
context of nuclear magnetic resonance. Larmor precession based imaging can be 
used to study the magnetic properties of solids at the level of magnetic 
domains. Neutron spin-echo is another technique borrowed from the world of 
nuclear magnetic resonance. It is used to study the dynamic properties of 
biologically important molecules. We closed the review with a section on the
unique ways in which neutron imaging is used to investigate macro-molecules
and detect elements in biological targets.

\bibliography{neutron}
\bibliographystyle{plain}
\end{document}
